Dissertation Proposal

### 2023-11-09 proposal.tex
grammar check

I plan to change the tutor such that it collects data to uncover these
unknown misconceptions and to confirm known ones.

@@

The preliminary tutor provides a ``Copy This Program'' button that
help students to run the program in DrRacket, where a SMoL
implementation is provided.

This approach is inconvenient both for the students and for me. For
students, they have to click, pastes and run, which is less convenient
than a single click. For me, because students run program in DrRacket
rather than ``within'' the tutor, I do not know whether they have run
the programs. This limit my ability to claim that the tutor is using
NM to help teaching.

@@

Another issue with the current design is that students might run a
program before giving an answer. I don't know yet how often this
happens. But given that students receive penalty (\Cref{s:penalty})
for giving wrong answers, they might do so when the program is overly
complicated. This problem is more of an issue after the switch to a
web-based notional machine because now the ``cost'' of running a
program is cheaper. My plan is to change the tutor so that it provides
a program-specific link \emph{only} after each interpreting task.

@@

The tutor is adding extra options in interpreting tasks in an ad hoc
way. Eventually, I plan to generate options by moving and/or copying
numbers in the current options. In the case of the example question in
\Cref{s:interpreting-tasks}, my plan would add \lstinline|12 12| and
\lstinline|0 12|. In summary, each MCQ has the following options:
\begin{itemize}
  \item the correct answer,
  \item wrong answers that correspond to known misconceptions,
  \item options generated by moving and/or copying numbers in the
        current options
  \item the ``Other'' option (which, once chosen, allows students to
        enter arbitrary answers)
\end{itemize}

@@

By the end of Fall 2023, I also want to improve the SMoL Tutor at
least on the following aspects:
\begin{enumerate}
  \item The tutor currently does not ask students to explain when they
        gave a wrong answer in an interpreting question. This limit my
        ability to find new misconceptions and to confirm that the
        tasks do detect the anticipated misconceptions. I plan to
        change the tutor so that it asks for an explanation in each
        interpreting task.
        % With student explanations, I check whether
        % choosing the wrong answer indicates holding the corresponding
        % misconception(s). I plan to eyeball the explanations to find
        % new misconceptions. This piece of data also helps me confirm
        % that anticipated wrong answers represent the expected
        % misconceptions.
  \item As I said in \Cref{s:effect-tutor}, for many of the
        misconceptions, the tutor has too few questions. On the other
        hand, there are many questions that do not produce useful data
        because their wrong answers do not uniquely point to a
        misconception. I plan to adjust the collection of interpreting
        questions such that every misconception has at least two
        questions that aim at correcting the misconception.
  \item The tutor is not randomizing the order of questions. This make
        it difficult to compare questions for one misconception --
        earlier questions might be difficult simply because the
        programs are somehow more difficult. I plan to randomize the
        order of interpreting tasks within each tutorial module.
  \item After randomizing question orders, I would need to check
        carefully the dependencies between explanations and to
        resolve these dependencies, if any.
  \item The tutor is using many variable names, some are separated by
        dash (e.g., \lstinline|get-x|). Students might misread these
        variables. For example, some students might misread
        \lstinline|(get-x)| (a function named \lstinline|get-x| is
        called with no argument) as \lstinline|(get x)| (a function
        named \lstinline|get| is called with one argument
        \lstinline|x|). These misreading add noise to the data. So I plan
        to use consistent and simple variable names in all programs.
\end{enumerate}

@@

I am officially taking a personal leave to do military service in my
country dring Spring 2024, so I do not plan to make more progress
other than data collection.

During Fall 2024, I plan to collect a new round of data from Brown. I
expect these datasets and the datasets from Spring 2024 would give a
more rigorous evaluation of the effectiveness of SMoL Tutor. The
improvements to be made in Fall 2023 should resolve many of the
threats to validity. In this semester I also want to write a new paper
about the SMoL Tutor.

>>>>

I plan to change the tutor such that it collects data to uncover these
unknown misconceptions and confirm known ones.

@@

The preliminary tutor provides a ``Copy This Program'' button that
helps students to run the program in DrRacket, where a SMoL
implementation is provided.

This approach is inconvenient both for the students and for me. For
students, they have to click, paste, and run, which is less convenient
than a single click. For me, because students run program in DrRacket
rather than ``within'' the tutor, I do not know whether they have run
the programs. This limits my ability to claim that the tutor is using
NM to help teach.

@@

Another issue with the current design is that students might run a
program before giving an answer. I don't know yet how often this
happens. But, given that students receive a penalty (\Cref{s:penalty})
for giving wrong answers, they might do so when the program is overly
complicated. This problem is more of an issue after the switch to a
web-based notional machine because now the ``cost'' of running a
program is cheaper. My plan is to change the tutor so that it provides
a program-specific link \emph{only} after each interpreting task.

@@

The tutor is adding extra options to interpreting tasks in an ad hoc
way. Eventually, I plan to generate options by moving and/or copying
numbers in the current options. In the case of the example question in
\Cref{s:interpreting-tasks}, my plan would add \lstinline|12 12| and
\lstinline|0 12|. In summary, each MCQ has the following options:
\begin{itemize}
  \item the correct answer,
  \item wrong answers that correspond to known misconceptions,
  \item options generated by moving and/or copying numbers in the
        current options
  \item the ``Other'' option (which, once chosen, allows students to
        enter arbitrary answers)
\end{itemize}

@@

By the end of Fall 2023, I also want to improve the SMoL Tutor at
least on the following aspects:
\begin{enumerate}
  \item The tutor currently does not ask students to explain when they
        gave a wrong answer in an interpreting question. This limits my
        ability to find new misconceptions and to confirm that the
        tasks do detect the anticipated misconceptions. I plan to
        change the tutor so that it asks for an explanation in each
        interpreting task.
        % With student explanations, I check whether
        % choosing the wrong answer indicates holding the corresponding
        % misconception(s). I plan to eyeball the explanations to find
        % new misconceptions. This piece of data also helps me confirm
        % that anticipated wrong answers represent the expected
        % misconceptions.
  \item As I said in \Cref{s:effect-tutor}, for many of the
        misconceptions, the tutor has too few questions. On the other
        hand, there are many questions that do not produce useful data
        because their wrong answers do not uniquely point to a
        misconception. I plan to adjust the collection of interpreting
        questions so that every misconception has at least two
        questions that aim to correct the misconception.
  \item The tutor is not randomizing the order of questions. This makes
        it difficult to compare questions for one misconception --
        earlier questions might be difficult simply because the
        programs are somehow more difficult. I plan to randomize the
        order of interpreting tasks within each tutorial module.
  \item After randomizing question orders, I would need to check
        carefully the dependencies between explanations and
        resolve these dependencies, if any.
  \item The tutor uses many variable names, some are separated by dashes
        (e.g., \lstinline|get-x|). Students might misread these
        variables. For example, some students might misread
        \lstinline|(get-x)| (a function named \lstinline|get-x| is
        called with no argument) as \lstinline|(get x)| (a function
        named \lstinline|get| is called with one argument
        \lstinline|x|). These misreadings add noise to the data. So I
        plan to use consistent and simple variable names in all
        programs.
\end{enumerate}

@@

I am officially taking a personal leave to do military service in my
country in Spring 2024, so I do not plan to make more progress
other than data collection.

During Fall 2024, I plan to collect a new round of data from Brown, analyze
data that I collected in Spring 2024, and hopefully write a paper about the
updates.
